Streaking and Clutching in Round-Based First Person Shooter Esports

### 2022-07-17 presentation.qmd
idk if i can stop the reformatting of bullets, tried looking at momentum from 2nd paper

---
title: "Streaking and Clutching in Round-Based First Person Shooter Esports"
author: "Tony ElHabr"
format:
  revealjs:
    incremental: false
    theme: [dark, theme.scss]
execute: 
  include: false
  echo: false
  eval: false
---

@@

```{r}
#| label: cod_heatmap_prep
n_rounds_side <- rounds |> 
  group_by(pre_cumu_w, pre_cumu_l, is_offense) |> 
  summarize(
    n = n(),
    across(c(win_round, win_series), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_round_prop = win_round / n,
    win_series_prop = win_series / n
  )

o_win_prop <- n_rounds_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!o_win_prop_const
  )

@@

@@

heatmap_seq_labeller <- function(.data, .label) {
  sprintf(
    '%s\n(%s/%s)', 
    scales::percent(.data[[.label]], accuracy = 1), 
    scales::comma(.data$win_round), 
    scales::comma(.data$n)
  )
}

heatmap_text_layer <- function(..., .op, .data, .label, .color, .threshold, .labeller) {
  list(
    ...,
    geom_text(
      data = .data |> filter(`.op`(.data[[.label]], .threshold)),
      # family = font,
      color = .color,
      size = 11 / .pt,
      fontface = 'bold',
      aes(
        x = pre_cumu_w + 0.5, 
        y = pre_cumu_l + 0.5,
        label = .labeller(.data = .data, .label = .label)
      )
    )
  )
}

@@

```{r}
#| label: p_o_round_win_prop
o_round_win_text_layer <- function(..., .emphasize = 'low') {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'win_round_prop',
      .color = .color,
      .threshold = 0.52,
      .labeller = heatmap_seq_labeller
    )
  )
}

@@

## How do the empirical round results compare to expectation?

```{r}
#| label: p_o_round_win_prop_over_expected
heatmap_div_labeller <- function(.data, .label) {
  sprintf(
    '%+.0f\n(%s/%s)', 
    100 * .data[[.label]], 
    scales::comma(.data$win_round), 
    scales::comma(.data$n)
  )
}

@@

```{r}
#| label: p_o_series_win_prop
add_o_series_win_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'win_series_prop',
      .color = .color,
      .threshold = 0.52,
      .labeller = heatmap_seq_labeller
    )
  )
}

p_o_series_win_prop <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_series_prop') +
  heatmap_text_layers(.f = add_o_series_win_text_layer) +
  ggsci::scale_fill_material('deep-orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_o_series_win_prop
save_for_slide(p_o_series_win_prop)
```

@@

## How do the empirical series results compare to expectation?

@@

plot_round_vs_post_round_slope_chart <- function(.round) {
  o_win_rx_prop <- summarize_o_round_win_prop(.round)
  x1_label <- sprintf('Round %d<br/>o Win %%', .round)
  x2_label <- sprintf('Post-Round %d<br/>o Win %%', .round)

@@

p_series_outcome_prop_over_expected <- long_series_outcome_prop |> 
  ggplot() +
  aes(x = record, y = outcome_idx) +
  geom_tile(
    data = long_series_outcome_prop |> filter(outcome_idx > 1),
    show.legend = FALSE,
    # height = 0.8,
    aes(fill = value)
  ) +
  ggsci::scale_fill_material(
    palette = 'orange'
  ) +
  ggnewscale::new_scale_fill() +
  geom_tile(
    data = long_series_outcome_prop |> filter(outcome_idx == 1),
    show.legend = FALSE,
    height = 0.8,
    aes(fill = value)
  ) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_series_outcome_prop, 
      +max_diff_series_outcome_prop
    )
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 3),
    color = blackish_background,
    size = 11 / .pt,
    aes(label = sprintf('%.1f%%\n(%s)', 100 * value, scales::comma(n)))
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 2),
    color = blackish_background,
    size = 11 / .pt,
    aes(label = sprintf('%.1f%%', 100 * value))
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 1) |> filter(abs(value) != max(abs(value))),
    color = blackish_background,
    size = 11 / .pt,
    aes(label = sprintf('%+.1f%%', 100 * value))
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 1) |> filter(abs(value) == max(abs(value))),
    color = 'white',
    size = 11 / .pt,

@@

simulate_post_streak_prob <- function(sims = 1000, ...) {
  rerun(
    sims,
    get_post_streak_prob(...)
  ) |> 
    flatten_dbl() |> 
    mean(na.rm = TRUE)
}
# prob_n6_k2 <- simulate_post_streak_prob(sims = 10000, n = 6, k = 2, p = 0.5)
prob_n6_k3 <- simulate_post_streak_prob(sims = 10000, n = 6, k = 3, p = 0.5)

round_streaks <- rounds |> 
  filter(round <= 6) |> 
  group_by(year, sheet, series, team) |> 
  mutate(
    won_prior_round2 = lag(win_round, n = 2, default = NA),
    won_prior_round3 = lag(win_round, n = 3, default = NA)
  ) |> 
  ungroup() |>
  select(
    year,
    sheet,
    series,
    team,
    is_offense,
    win_round,
    won_prior_round,
    won_prior_round2,
    won_prior_round3
  )

round_streaks_after_b2b <- round_streaks |> 
  filter(!is.na(won_prior_round2)) |> 
  count(win_round, won_prior_round, won_prior_round2, sort = TRUE)

round_win_prop_after_b2b_wins <- round_streaks_after_b2b |> 
  filter(won_prior_round, won_prior_round2) |> 
  mutate(
    total = sum(n),
    prop = n / total
  ) |> 
  filter(win_round) |> 
  select(n, total, prop)

@@

>>>>

---
title: "Streaking and Clutching in Round-Based First Person Shooter Esports"
author: "Tony ElHabr"
format:
  revealjs:
    slide-number: true
    incremental: false
    logo: images/cdl_gray.png
    theme: [dark, theme.scss]
execute: 
  include: false
  echo: false
  eval: false
---

@@

```{r}
#| label: cod_heatmap_prep
round_and_series_win_prop_by_side <- rounds |> 
  group_by(pre_cumu_w, pre_cumu_l, is_offense) |> 
  summarize(
    n = n(),
    across(c(win_round, win_series), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_round_prop = win_round / n,
    win_series_prop = win_series / n
  )

o_win_prop <- round_and_series_win_prop_by_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!o_win_prop_const
  )

@@

# round_and_series_win_prop <- rounds |> 
#   group_by(pre_cumu_w, pre_cumu_l) |> 
#   summarize(
#     n = n(),
#     across(c(win_round, win_series), sum)
#   ) |> 
#   ungroup() |> 
#   mutate(
#     win_round_prop = win_round / n,
#     win_series_prop = win_series / n
#   )
# 
# pluck_round_win_prop <- function(w = 0, l = 0) {
#   round_and_series_win_prop |> 
#     filter(pre_cumu_w == w, pre_cumu_l == l) |> 
#     pull(win_round_prop)
# }
# 
# compute_conditional_probs <- function(
    #     p00 = 0.5,
#     p10 = p00,
#     p20 = p00,
#     p30 = p00,
#     p40 = p00,
#     p50 = p00,
#     p11 = p00, p21 = p00, p31 = p00, p41 = p00, p51 = p00,
#     p52 = p00, p53 = p00, p54 = p00
# ) {
#   pr6 <- p10 * p20 * p30 * p40 * p50
#   # pr7 <- p51 * (p10 * p20 * p30 * p40 * (1 - p50) + p10 * p20 * p30 * (1 - p40) * p41 + p10 * p20 * (1 - p30) * p41 + p10 * (1 - p20) * p31 * p41 + (1 - p10) * p21 * p31 * p41)
#   pr7 <- (
#     (p00 * p10 * p20 * p30 * p40 * (1 - p50) * p51) +   ## aaaaaba
#     (p00 * p10 * p20 * p30 * (1 - p40) * p41 * p51) + ## aaaabaa
#     (p00 * p10 * p20 * (1 - p30) * p31 * p41 * p51) + ## aaabaaa
#     (p00 * p10 * (1 - p20) * p21 * p31 * p41 * p51) + ## aabaaaa
#     (p00 * (1 - p10) * p11 * p21 * p31 * p41 * p51) + ## abaaaaa
#     ((1 - p00) * (1 - p10) * p11 * p21 * p31 * p41 * p51) + ## baaaaaa
#     
#     ((1 - p00) * p10 * p20 * p30 * p40 * (1 - p50) * p51) + ## bbbbbab
#     ((1 - p00) * p10 * p20 * p30 * (1 - p40) * p41 * p51) + ## bbbbabb
#     ((1 - p00) * p10 * p20 * (1 - p30) * p31 * p41 * p51) + ## bbbabbb
#     ((1 - p00) * p10 * (1 - p20) * p21 * p31 * p41 * p51) + ## bbabbbb
#     ((1 - p00) * (1 - p10) * (1 - p11) * p21 * p31 * p41 * p51) + ## babbbbb
#     (p00 * (1 - p10) * (1 - p11) * p21 * p31 * p41 * p51) ## abbbbbb
# )
#   
#   ## NOTE: pr8 and pr9 are broken. although we can use the simple subtraction of other probabilities for pr10 and pr11 like pr6 and pr7 in a 7 game series, they aren't valid for pr8 and pr9 in an 11 game series.
#   pr8 <- (1 - pr6 - pr7) * p52
#   pr9 <- (1 - pr6 - pr7 - pr8) * p53
#   pr10 <- (1 - pr6 - pr7 - pr8 - pr9) * p54
#   pr11 <- 1 - pr6 - pr7 - pr8 - pr9 - pr10
#   tibble(
#     n_rounds = 6L:11L,
#     p = c(pr6, pr7, pr8, pr9, pr10, pr11)
#   )
# }
# 
# ## pr6 and pr7 should match up with what we see later with `possibilities`
# conditional_series_probs <- compute_conditional_probs(
#   p00 = pluck_round_win_prop(0, 0),
#   p10 = pluck_round_win_prop(1, 0),
#   p20 = pluck_round_win_prop(2, 0),
#   p30 = pluck_round_win_prop(3, 0),
#   p40 = pluck_round_win_prop(4, 0),
#   p50 = pluck_round_win_prop(5, 0),
#   p11 = pluck_round_win_prop(1, 1),
#   p21 = pluck_round_win_prop(2, 1),
#   p31 = pluck_round_win_prop(3, 1),
#   p41 = pluck_round_win_prop(4, 1),
#   p51 = pluck_round_win_prop(5, 1),
#   p52 = pluck_round_win_prop(5, 2),
#   p53 = pluck_round_win_prop(5, 3),
#   p54 = pluck_round_win_prop(5, 4)
# )

@@

heatmap_seq_labeller <- function(.data, .label, .num) {
  sprintf(
    '%s\n(%s/%s)', 
    scales::percent(.data[[.label]], accuracy = 1), 
    scales::comma(.data[[.num]]), 
    scales::comma(.data$n)
  )
}

heatmap_text_layer <- function(..., .op, .data, .label, .num, .color, .threshold, .labeller) {
  list(
    ...,
    geom_text(
      data = .data |> filter(`.op`(.data[[.label]], .threshold)),
      # family = font,
      color = .color,
      size = 11 / .pt,
      fontface = 'bold',
      aes(
        x = pre_cumu_w + 0.5, 
        y = pre_cumu_l + 0.5,
        label = .labeller(.data = .data, .label = .label, .num)
      )
    )
  )
}

@@

```{r}
#| label: p_o_round_win_prop
o_round_win_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'win_round_prop',
      .num = 'win_round',
      .color = .color,
      .threshold = 0.52,
      .labeller = heatmap_seq_labeller
    )
  )
}

@@

```{r}
#| label: p_o_round_win_prop_over_expected
# I have no idea why I need to add .num here... keeps erroring otherwise
heatmap_div_labeller <- function(.data, .label, .num) {
  sprintf(
    '%+.1f\n(%s)', 
    100 * .data[[.label]], 
    scales::comma(.data$n)
  )
}

@@

```{r}
#| label: p_o_series_win_prop
add_o_series_win_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'win_series_prop',
      .num = 'win_series',
      .color = .color,
      .threshold = 0.8,
      .labeller = heatmap_seq_labeller
    )
  )
}

p_o_series_win_prop <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_series_prop') +
  heatmap_text_layers(.f = add_o_series_win_text_layer) +
  ggsci::scale_fill_material('orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_o_series_win_prop
save_for_slide(p_o_series_win_prop)
```

@@

@@

plot_round_vs_post_round_slope_chart <- function(.round) {
  o_win_rx_prop <- summarize_o_round_win_prop(.round)
  x1_label <- sprintf('Round %d<br/>offensive win %%', .round)
  x2_label <- sprintf('Post-round %d<br/>offensive win %%', .round)

@@

p_series_outcome_prop_over_expected <- long_series_outcome_prop |> 
  ggplot() +
  aes(x = record, y = outcome_idx) +
  geom_tile(
    data = long_series_outcome_prop |> filter(outcome_idx > 1),
    show.legend = FALSE,
    # height = 0.8,
    aes(fill = value)
  ) +
  ggsci::scale_fill_material(
    palette = 'deep-orange'
  ) +
  ggnewscale::new_scale_fill() +
  geom_tile(
    data = long_series_outcome_prop |> filter(outcome_idx == 1),
    show.legend = FALSE,
    height = 0.8,
    aes(fill = value)
  ) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_series_outcome_prop, 
      +max_diff_series_outcome_prop
    )
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 3),
    color = blackish_background,
    size = 11 / .pt,
    aes(label = sprintf('%.1f%%\n(%s)', 100 * value, scales::comma(n)))
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 2),
    color = blackish_background,
    size = 11 / .pt,
    aes(label = sprintf('%.1f%%', 100 * value))
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 1) |> filter(abs(value) != max(abs(value))),
    color = blackish_background,
    size = 11 / .pt,
    aes(label = sprintf('%+.1f%%', 100 * value))
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 1) |> filter(abs(value) == max(abs(value))),
    color = 'white',
    size = 11 / .pt,

@@

simulate_post_streak_prob <- function(sims = 1000, ...) {
  rerun(
    sims,
    get_post_streak_prob(...)
  ) |> 
    flatten_dbl() |> 
    mean(na.rm = TRUE)
}
# prob_n6_k2 <- simulate_post_streak_prob(sims = 10000, n = 6, k = 2, p = 0.5)
prob_n6_k3 <- simulate_post_streak_prob(sims = 10000, n = 6, k = 3, p = 0.5)
prob_n11_k3 <- simulate_post_streak_prob(sims = 10000, n = 8, k = 3, p = 0.5)
prob_n11_k3 <- simulate_post_streak_prob(sims = 10000, n = 11, k = 3, p = 0.5)
prob_n10_k1 <- simulate_post_streak_prob(sims = 10000, n = 10, k = 1, p = 0.5)

round_streaks <- rounds |> 
  # filter(round <= 6) |> 
  group_by(year, sheet, series, team) |> 
  mutate(
    won_prior_round2 = lag(win_round, n = 2, default = NA),
    won_prior_round3 = lag(win_round, n = 3, default = NA)
  ) |> 
  ungroup() |>
  select(
    year,
    sheet,
    series,
    team,
    is_offense,
    round,
    win_round,
    won_prior_round,
    won_prior_round2,
    won_prior_round3
  )

@@

```{r}
#| label: summarize_and_compare_win_prop_after_wins
## https://content.iospress.com/articles/journal-of-sports-analytics/jsa200422#app2 suggests that we should look at non-overlapping streaks.
## "To avoid this problem, we examine only the frequency of winning game three after having won games one and two and the frequency of winning game six after having won games four and five". They sort of deduced that win %s > 50% in the third game might constitute momentum, but didn't make strong statements.
summarize_win_prop_after_x_wins <- function(round, ...) {
  round_streaks |> 
    filter(round == !!round) |> 
    count(win_round, won_prior_round, won_prior_round2, ..., sort = TRUE) |> 
    relocate(n) |> 
    drop_na(last_col()) |> 
    filter(won_prior_round, won_prior_round2, ...) |> 
    mutate(
      total = sum(n),
      prop = n / total
    ) |> 
    filter(win_round) |> 
    transmute(round = !!round, n, total, prop)
}

summarize_win_prop_after_b2b_wins <- summarize_win_prop_after_x_wins
summarize_win_prop_after_b2b2b_wins <- function(round) {
  summarize_win_prop_after_x_wins(round = round, won_prior_round3)
}

summarize_and_compare_win_prop_after_wins <- function(ri, add, f) {
  ra <- f(ri)
  rb <- f(ri + add)
  bind_rows(ra, rb)
}

summarize_and_compare_win_prop_after_b2b_wins <- partial(
  summarize_and_compare_win_prop_after_wins,
  add = 3,
  f = summarize_win_prop_after_b2b_wins,
  ... = 
)

summarize_and_compare_win_prop_after_b2b2b_wins <- partial(
  summarize_and_compare_win_prop_after_wins,
  add = 4,
  f = summarize_win_prop_after_b2b2b_wins,
  ... = 
)

win_prop_after_b2b_wins <- seq(3, 8, by = 1) |> 
  map_dfr(
    summarize_and_compare_win_prop_after_b2b_wins,
    .id = 'pair_idx'
  )

win_prop_after_b2b2b_wins <- seq(4, 7, by = 1) |> 
  map_dfr(
    summarize_and_compare_win_prop_after_b2b2b_wins,
    .id = 'pair_idx'
  )
```
